---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---

##Loading and preprocessing the data

1. Load the data

```{r, echo = TRUE}
url <- "https://raw.githubusercontent.com/joneschoi/RepData_PeerAssessment1/master/activity.zip"
f <- file.path(getwd(), "activity.zip")
download.file(url, f)
unzip("activity.zip")
act <- read.csv("activity.csv")
head(act)
```

##What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r, echo = TRUE}
ActDate <- aggregate( steps~date, act, sum)
barplot(ActDate$steps, name = as.Date(ActDate$date), xlab ="Date", ylab="Step", main="The Total Number of Steps Taken (Each Day)")
```

2. Calculate and report the mean and median total number of steps taken per day

```{r, echo = TRUE}
x <- act[act$steps != 0 ,]
print(aggregate( steps~date, x, FUN = function(x) c(mean = mean(x), median = median(x))), row.names = FALSE)
```

##What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo = TRUE}
ActInt <- aggregate( steps~interval, act, mean)
plot(ActInt$interval, ActInt$steps, type = "l", main = "The Average Number of Steps Taken (Interval)", xlab = "Interval", ylab = "Steps" )
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo = TRUE}
print(ActInt[ActInt$steps == max(ActInt$steps),], row.names = FALSE)
```

##Imputing missing values

1. Calculate and report the total number of missing values in the dataset 

```{r, echo = TRUE}
sum(is.na(act$steps))
```

2. Devise a strategy for filling in all of the missing values in the dataset. This use the mean for that 5-minute interval.

```{r, echo = TRUE}
IntAss <- aggregate( steps~interval, act, mean)
head(IntAss)
```

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo = TRUE}
Actna <- act[is.na(act$steps) == TRUE,]
Actnay <- merge(x=Actna,y=IntAss,by="interval", all.x = TRUE)
Actnay <- Actnay[,c(4,3,1)]
names(Actnay)[1] = "steps"
Actnan <- act[is.na(act$steps) == FALSE,]
ActAdd <- rbind(Actnan,Actnay)
head(ActAdd)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. We know that the first report day and the last report day is fill up the same mean.

```{r, echo = TRUE}
Actmem <- aggregate( steps~date, ActAdd, mean)
barplot(Actmem$steps, name = as.Date(Actmem$date), xlab ="Date", ylab="Step", main="The Mean of Steps Taken (Each Day)")

Actmed <- ActAdd[ActAdd$steps != 0 ,]
Actmed <- aggregate( steps~date, Actmed, median)
barplot(Actmed$steps, name = as.Date(Actmed$date), xlab ="Date", ylab="Step", main="The median of Steps Taken (Each Day)")
```

##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r, echo = TRUE, results="hide"}
Sys.setlocale("LC_TIME", "English")
```

```{r, echo = TRUE}
ActAdd$wkdy <- weekdays(as.Date(ActAdd$date))
ActEnd <- ActAdd[ActAdd$wkdy %in% c("Saturday","Sunday"),]
ActDay <- ActAdd[!(ActAdd$wkdy %in% c("Saturday","Sunday")),]
```

```{r, echo = TRUE}
ActEnd$wkdy <- "Weekend"
head(ActEnd)
```

```{r, echo = TRUE}
ActDay$wkdy <- "Weekday"
head(ActDay)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r, echo = TRUE}
ActEnda <- aggregate( steps~interval, ActEnd, mean)
ActDaya <- aggregate( steps~interval, ActDay, mean)
par(mfrow=c(2,1))
plot(ActEnda$interval, ActEnda$steps, type = "l", main = "The Average Number of Steps Taken (Weekend)", xlab = "Interval", ylab = "Steps" )
plot(ActDaya$interval, ActDaya$steps, type = "l", main = "The Average Number of Steps Taken (Weekday)", xlab = "Interval", ylab = "Steps" )
```
